---
title: "Data Clean"
author: "Jingjian Wang"
date: '2022-11-28'
output: html_document
---

library
```{r}
pacman::p_load(haven, tidyverse, dplyr)
```

import data from 2016 to 2021

```{r}
D21<-read_xpt("/Users/jingjianwang/Documents/5245.01 Advanced Data Analysis/LLCP2021.XPT ",
col_select = c("PREGNANT", "RENTHOM1", "GENHLTH", "_INCOMG1", "_EDUCAG", "_MENT14D", "_AGE80", "_AGE_G", "_IMPRACE", "EMPLOY1"),skip = 0,n_max = Inf,.name_repair = "unique")
D20<-read_xpt("/Users/jingjianwang/Documents/5245.01 Advanced Data Analysis/LLCP2020.XPT ",
col_select = c("PREGNANT", "RENTHOM1", "GENHLTH", "_INCOMG", "_EDUCAG", "_MENT14D", "_AGE80", "_AGE_G", "_IMPRACE", "EMPLOY1"),skip = 0,n_max = Inf,.name_repair = "unique")
D19<-read_xpt("/Users/jingjianwang/Documents/5245.01 Advanced Data Analysis/LLCP2019.XPT ",
col_select = c("PREGNANT", "RENTHOM1", "GENHLTH", "_INCOMG", "_EDUCAG", "_MENT14D", "_AGE80", "_AGE_G", "_IMPRACE", "EMPLOY1"),skip = 0,n_max = Inf,.name_repair = "unique")
D18<-read_xpt("/Users/jingjianwang/Documents/5245.01 Advanced Data Analysis/LLCP2018.XPT ",
col_select = c("PREGNANT", "RENTHOM1", "GENHLTH", "_INCOMG", "_EDUCAG", "_MENT14D", "_AGE80", "_AGE_G", "_IMPRACE", "EMPLOY1"),skip = 0,n_max = Inf,.name_repair = "unique")
D17<-read_xpt("/Users/jingjianwang/Documents/5245.01 Advanced Data Analysis/LLCP2017.XPT ",
col_select = c("PREGNANT", "RENTHOM1", "GENHLTH", "_INCOMG", "_EDUCAG", "_MENT14D", "_AGE80", "_AGE_G", "_IMPRACE", "EMPLOY1"),skip = 0,n_max = Inf,.name_repair = "unique")
D16<-read_xpt("/Users/jingjianwang/Documents/5245.01 Advanced Data Analysis/LLCP2016.XPT ",
col_select = c("PREGNANT", "RENTHOM1", "GENHLTH", "_INCOMG", "_EDUCAG", "_MENT14D", "_AGE80", "_AGE_G", "_RACE", "EMPLOY1"),skip = 0,n_max = Inf,.name_repair = "unique")
```
In variable Mental Health Status, those with more than half of the days in a month feeling not good are regarded as with 'Bad' mental health status.
In variable Employment, student is regarded as being employed.
In variable Age, the age was recoded as '18-34' years old and more than 34 years old.
In variable Education, education level was recoded as 'less than high school', 'high school', and 'college or technical school'.

```{r}
D21.little <- D21 %>%
  filter(PREGNANT == 1 & `_INCOMG1` %in% c('4','5','6','7')) %>%
  mutate(RENTHOM = recode_factor(RENTHOM1, 
                                '1' = 'No',
                                '2' = 'Yes',
                                '3' = NA_character_,
                                '7' = NA_character_,
                                '9' = NA_character_)) %>%
  mutate(GHLTH = recode_factor(GENHLTH, 
                                '1' = 'Good',
                                '2' = 'Good',
                                '3' = 'Good',
                                '4' = 'Bad',
                                '5' = 'Bad',
                                '7' = NA_character_,
                                '9' = NA_character_)) %>%
  mutate(MHLTH = recode_factor(`_MENT14D`, 
                                '1' = 'Good',
                                '2' = 'Good',
                                '3' = 'Bad',
                                '9' = NA_character_)) %>%
  mutate(EMPLOY = recode_factor(EMPLOY1, 
                                '1' = 'Yes',
                                '2' = 'Yes',
                                '3' = 'No',
                                '4' = 'No',
                                '5' = 'No',
                                '6' = 'Yes',
                                '7' = 'No',
                                '8' = 'No',
                                '9' = NA_character_)) %>%
  mutate(Race = recode_factor(`_IMPRACE`,
                                '1' = 'White, NH',
                                '2' = 'Black, NH',
                                '3' = 'Asian, NH',
                                '4' = 'American Indian/Alaska Native,NH',
                                '5' = 'Hispanic',
                                '6' = 'Other, NH')) %>%
  mutate(Age = `_AGE80`) %>%
  mutate(AgeG = recode_factor(`_AGE_G`,
                                '1' = '18-34',
                                '2' = '18-34',
                                '3' = '34+',
                                '4' = '34+',
                                '5' = '34+',
                                '6' = '34+')) %>%
  mutate(EDU = recode_factor(`_EDUCAG`,
                                '1' = 'Less than High School',
                                '2' = 'High School',
                                '3' = 'College or Technical school',
                                '4' = 'College or Technical school',
                                '9' = NA_character_)) %>%
  mutate(YEAR = 2021) %>%
  select("RENTHOM", "GHLTH", "MHLTH", "EMPLOY", "Race", "Age", "AgeG", "EDU", "YEAR")
summary(D21.little)
```

Before 2021, the variable `_INCOMG1` is not available, and `_INCOMG` is used here.
```{r}
D20.little <- D20 %>%
  filter(PREGNANT == 1 & `_INCOMG` %in% c('4','5')) %>%
  mutate(RENTHOM = recode_factor(RENTHOM1, 
                                '1' = 'No',
                                '2' = 'Yes',
                                '3' = NA_character_,
                                '7' = NA_character_,
                                '9' = NA_character_)) %>%
  mutate(GHLTH = recode_factor(GENHLTH, 
                                '1' = 'Good',
                                '2' = 'Good',
                                '3' = 'Good',
                                '4' = 'Bad',
                                '5' = 'Bad',
                                '7' = NA_character_,
                                '9' = NA_character_)) %>%
  mutate(MHLTH = recode_factor(`_MENT14D`, 
                                '1' = 'Good',
                                '2' = 'Good',
                                '3' = 'Bad',
                                '9' = NA_character_)) %>%
  mutate(EMPLOY = recode_factor(EMPLOY1, 
                                '1' = 'Yes',
                                '2' = 'Yes',
                                '3' = 'No',
                                '4' = 'No',
                                '5' = 'No',
                                '6' = 'Yes',
                                '7' = 'No',
                                '8' = 'No',
                                '9' = NA_character_)) %>%
  mutate(Race = recode_factor(`_IMPRACE`,
                                '1' = 'White, NH',
                                '2' = 'Black, NH',
                                '3' = 'Asian, NH',
                                '4' = 'American Indian/Alaska Native,NH',
                                '5' = 'Hispanic',
                                '6' = 'Other, NH')) %>%
  mutate(Age = `_AGE80`) %>%
  mutate(AgeG = recode_factor(`_AGE_G`,
                                '1' = '18-34',
                                '2' = '18-34',
                                '3' = '34+',
                                '4' = '34+',
                                '5' = '34+',
                                '6' = '34+')) %>%
  mutate(EDU = recode_factor(`_EDUCAG`,
                                '1' = 'Less than High School',
                                '2' = 'High School',
                                '3' = 'College or Technical school',
                                '4' = 'College or Technical school',
                                '9' = NA_character_)) %>%
  mutate(YEAR = 2020) %>%
  select("RENTHOM", "GHLTH", "MHLTH", "EMPLOY", "Race", "Age", "AgeG", "EDU", "YEAR")
summary(D20.little)
```


```{r}
D19.little <- D19 %>%
  filter(PREGNANT == 1 & `_INCOMG` %in% c('4','5')) %>%
  mutate(RENTHOM = recode_factor(RENTHOM1, 
                                '1' = 'No',
                                '2' = 'Yes',
                                '3' = NA_character_,
                                '7' = NA_character_,
                                '9' = NA_character_)) %>%
  mutate(GHLTH = recode_factor(GENHLTH, 
                                '1' = 'Good',
                                '2' = 'Good',
                                '3' = 'Good',
                                '4' = 'Bad',
                                '5' = 'Bad',
                                '7' = NA_character_,
                                '9' = NA_character_)) %>%
  mutate(MHLTH = recode_factor(`_MENT14D`, 
                                '1' = 'Good',
                                '2' = 'Good',
                                '3' = 'Bad',
                                '9' = NA_character_)) %>%
  mutate(EMPLOY = recode_factor(EMPLOY1, 
                                '1' = 'Yes',
                                '2' = 'Yes',
                                '3' = 'No',
                                '4' = 'No',
                                '5' = 'No',
                                '6' = 'Yes',
                                '7' = 'No',
                                '8' = 'No',
                                '9' = NA_character_)) %>%
  mutate(Race = recode_factor(`_IMPRACE`,
                                '1' = 'White, NH',
                                '2' = 'Black, NH',
                                '3' = 'Asian, NH',
                                '4' = 'American Indian/Alaska Native,NH',
                                '5' = 'Hispanic',
                                '6' = 'Other, NH')) %>%
  mutate(Age = `_AGE80`) %>%
  mutate(AgeG = recode_factor(`_AGE_G`,
                                '1' = '18-34',
                                '2' = '18-34',
                                '3' = '34+',
                                '4' = '34+',
                                '5' = '34+',
                                '6' = '34+')) %>%
  mutate(EDU = recode_factor(`_EDUCAG`,
                                '1' = 'Less than High School',
                                '2' = 'High School',
                                '3' = 'College or Technical school',
                                '4' = 'College or Technical school',
                                '9' = NA_character_)) %>%
  mutate(YEAR = 2019) %>%
  select("RENTHOM", "GHLTH", "MHLTH", "EMPLOY", "Race", "Age", "AgeG", "EDU", "YEAR") 
summary(D19.little)
```



```{r}
D18.little <- D18 %>%
  filter(PREGNANT == 1 & `_INCOMG` %in% c('4','5')) %>%
  mutate(RENTHOM = recode_factor(RENTHOM1, 
                                '1' = 'No',
                                '2' = 'Yes',
                                '3' = NA_character_,
                                '7' = NA_character_,
                                '9' = NA_character_)) %>%
  mutate(GHLTH = recode_factor(GENHLTH, 
                                '1' = 'Good',
                                '2' = 'Good',
                                '3' = 'Good',
                                '4' = 'Bad',
                                '5' = 'Bad',
                                '7' = NA_character_,
                                '9' = NA_character_)) %>%
  mutate(MHLTH = recode_factor(`_MENT14D`, 
                                '1' = 'Good',
                                '2' = 'Good',
                                '3' = 'Bad',
                                '9' = NA_character_)) %>%
  mutate(EMPLOY = recode_factor(EMPLOY1, 
                                '1' = 'Yes',
                                '2' = 'Yes',
                                '3' = 'No',
                                '4' = 'No',
                                '5' = 'No',
                                '6' = 'Yes',
                                '7' = 'No',
                                '8' = 'No',
                                '9' = NA_character_)) %>%
  mutate(Race = recode_factor(`_IMPRACE`,
                                '1' = 'White, NH',
                                '2' = 'Black, NH',
                                '3' = 'Asian, NH',
                                '4' = 'American Indian/Alaska Native,NH',
                                '5' = 'Hispanic',
                                '6' = 'Other, NH')) %>%
  mutate(Age = `_AGE80`) %>%
  mutate(AgeG = recode_factor(`_AGE_G`,
                                '1' = '18-34',
                                '2' = '18-34',
                                '3' = '34+',
                                '4' = '34+',
                                '5' = '34+',
                                '6' = '34+')) %>%
  mutate(EDU = recode_factor(`_EDUCAG`,
                                '1' = 'Less than High School',
                                '2' = 'High School',
                                '3' = 'College or Technical school',
                                '4' = 'College or Technical school',
                                '9' = NA_character_)) %>%
  mutate(YEAR = 2018) %>%
  select("RENTHOM", "GHLTH", "MHLTH", "EMPLOY", "Race", "Age", "AgeG", "EDU", "YEAR") 
summary(D18.little)
```



```{r}
D17.little <- D17 %>%
  filter(PREGNANT == 1 & `_INCOMG` %in% c('4','5')) %>%
  mutate(RENTHOM = recode_factor(RENTHOM1, 
                                '1' = 'No',
                                '2' = 'Yes',
                                '3' = NA_character_,
                                '7' = NA_character_,
                                '9' = NA_character_)) %>%
  mutate(GHLTH = recode_factor(GENHLTH, 
                                '1' = 'Good',
                                '2' = 'Good',
                                '3' = 'Good',
                                '4' = 'Bad',
                                '5' = 'Bad',
                                '7' = NA_character_,
                                '9' = NA_character_)) %>%
  mutate(MHLTH = recode_factor(`_MENT14D`, 
                                '1' = 'Good',
                                '2' = 'Good',
                                '3' = 'Bad',
                                '9' = NA_character_)) %>%
  mutate(EMPLOY = recode_factor(EMPLOY1, 
                                '1' = 'Yes',
                                '2' = 'Yes',
                                '3' = 'No',
                                '4' = 'No',
                                '5' = 'No',
                                '6' = 'Yes',
                                '7' = 'No',
                                '8' = 'No',
                                '9' = NA_character_)) %>%
  mutate(Race = recode_factor(`_IMPRACE`,
                                '1' = 'White, NH',
                                '2' = 'Black, NH',
                                '3' = 'Asian, NH',
                                '4' = 'American Indian/Alaska Native,NH',
                                '5' = 'Hispanic',
                                '6' = 'Other, NH')) %>%
  mutate(Age = `_AGE80`) %>%
  mutate(AgeG = recode_factor(`_AGE_G`,
                                '1' = '18-34',
                                '2' = '18-34',
                                '3' = '34+',
                                '4' = '34+',
                                '5' = '34+',
                                '6' = '34+')) %>%
  mutate(EDU = recode_factor(`_EDUCAG`,
                                '1' = 'Less than High School',
                                '2' = 'High School',
                                '3' = 'College or Technical school',
                                '4' = 'College or Technical school',
                                '9' = NA_character_)) %>%
  mutate(YEAR = 2017) %>%
  select("RENTHOM", "GHLTH", "MHLTH", "EMPLOY", "Race", "Age", "AgeG", "EDU", "YEAR") 
summary(D17.little)
```

In 2016, variable `_IMPRACE` is not available and `_RACE` is used here.

```{r}
D16.little <- D16 %>%
  filter(PREGNANT == 1 & `_INCOMG` %in% c('4','5')) %>%
  mutate(RENTHOM = recode_factor(RENTHOM1, 
                                '1' = 'No',
                                '2' = 'Yes',
                                '3' = NA_character_,
                                '7' = NA_character_,
                                '9' = NA_character_)) %>%
  mutate(GHLTH = recode_factor(GENHLTH, 
                                '1' = 'Good',
                                '2' = 'Good',
                                '3' = 'Good',
                                '4' = 'Bad',
                                '5' = 'Bad',
                                '7' = NA_character_,
                                '9' = NA_character_)) %>%
  mutate(MHLTH = recode_factor(`_MENT14D`, 
                                '1' = 'Good',
                                '2' = 'Good',
                                '3' = 'Bad',
                                '9' = NA_character_)) %>%
  mutate(EMPLOY = recode_factor(EMPLOY1, 
                                '1' = 'Yes',
                                '2' = 'Yes',
                                '3' = 'No',
                                '4' = 'No',
                                '5' = 'No',
                                '6' = 'Yes',
                                '7' = 'No',
                                '8' = 'No',
                                '9' = NA_character_)) %>%
  mutate(Race = recode_factor(`_RACE`,
                                '1' = 'White, NH',
                                '2' = 'Black, NH',
                                '4' = 'Asian, NH',
                                '3' = 'American Indian/Alaska Native,NH',
                                '8' = 'Hispanic',
                                '5' = 'Other, NH',
                                '6' = 'Other, NH',
                                '7' = 'Other, NH',
                                '9' = NA_character_)) %>%
  mutate(Age = `_AGE80`) %>%
  mutate(AgeG = recode_factor(`_AGE_G`,
                                '1' = '18-34',
                                '2' = '18-34',
                                '3' = '34+',
                                '4' = '34+',
                                '5' = '34+',
                                '6' = '34+')) %>%
  mutate(EDU = recode_factor(`_EDUCAG`,
                                '1' = 'Less than High School',
                                '2' = 'High School',
                                '3' = 'College or Technical school',
                                '4' = 'College or Technical school',
                                '9' = NA_character_)) %>%
  mutate(YEAR = 2016) %>%
  select("RENTHOM", "GHLTH", "MHLTH", "EMPLOY", "Race", "Age", "AgeG", "EDU", "YEAR")
summary(D16.little)
```


combine the datasets together and save
```{r}
D16_21 <- rbind(D16.little,D17.little,D18.little,D19.little,D20.little,D21.little) %>%
  drop_na()
write_xpt(D16_21,"/Users/jingjianwang/Documents/5245.01 Advanced Data Analysis/Final/LLCP1621.XPT")
```
